Comparison of approaches to using machine learning models to increase the efficiency of the global search algorithm for solving multicriterial problems

### 2024-08-09 2024/OPTIMA/OPTIMA_2024_MCO.tex
fix some references
correct one sentence
fix the references

% This is samplepaper.tex, a sample chapter demonstrating the
% LLNCS macro package for Springer Computer Science proceedings;
% Version 2.21 of 2022/01/12
%
\documentclass[runningheads]{llncs}
%
\usepackage[T1]{fontenc}
% T1 fonts will be used to generate the final print and online PDFs,
% so please use T1 fonts in your manuscript whenever possible.
% Other font encondings may result in incorrect characters.
%
\usepackage{multirow}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{cite}
\usepackage{hyperref}

\usepackage[misc,geometry]{ifsym}

@@

There are various ways to construct the function $F(\lambda, y)$ \cite{Miettinen1999,Ehrgott2005,GergelKozinov2020,Marler2004}. In our study we utilized the maximum convolution 
\begin{equation}
\label{eq:05}
F(\lambda, y) = \max_{1 \leq i \leq s} \left\{\lambda_i f_i (y)\right\}, y \in D, \sum_{i=1}^s {\lambda_i} = 1, 0 < \lambda_i \leq 1.
\end{equation}

@@

One of the ways how this can be done is connected with the modification of the starting iterations of GSA. Namely, if $P \geq 1$ problems (\ref{eq:07}) have been already solved and the matrix (\ref{eq:15}) has been formed, then as initial trial coordinates, the points $x^q$, $0 \leq q \leq K$, can be used with values $z^q = F(\lambda,y^q)$. Moreover, the computation of the convolution (\ref{eq:05}) does not require the new calculation of criteria values at these points, because these values are taken from the matrix (\ref{eq:15}). After that, one can go to the basic iteration Rules 1-5. Such the scheme reduces significantly the number of criteria computations when jointly solving the family of problems (\ref{eq:04}) with convolution (\ref{eq:05}) for different weight coefficients $\lambda$ \cite{ML_MCO_2023,Gergel2018,GergelKozinov2020}.

@@

For computation of $R_{ps}(i)$ we propose to apply the machine learning methods. It should be noted that in classical combinations of machine learning and optimization methods, the latter are used as an auxiliary internal techniques that improve the machine learning functioning. In our approach, on the contrary, machine learning methods are a part of the optimization algorithms enhancing their work.

>>>>

% This is samplepaper.tex, a sample chapter demonstrating the
% LLNCS macro package for Springer Computer Science proceedings;
% Version 2.21 of 2022/01/12
%
\documentclass[runningheads]{llncs}
%
\usepackage[T1]{fontenc}
% T1 fonts will be used to generate the final print and online PDFs,
% so please use T1 fonts in your manuscript whenever possible.
% Other font encondings may result in incorrect characters.
%
\usepackage{multirow}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{cite}
\usepackage{hyperref}
\usepackage{color}
\usepackage[misc,geometry]{ifsym}

@@

There are various ways to construct the function $F(\lambda, y)$ \cite{Miettinen1999,Ehrgott2005,GergelKozinov2020,Marler2004}. In our study we utilized the maximum convolution 
\begin{equation}
\label{eq:05}
F(\lambda, y) = \max_{1 \leq i \leq s} \left\{\lambda_i f_i (y)\right\}, y \in D, \sum_{i=1}^s {\lambda_i} = 1, 0 < \lambda_i < 1.
\end{equation}

@@

One of the ways how this can be done is connected with the modification of the starting iterations of GSA. Namely, if $P \geq 1$ problems (\ref{eq:07}) have been already solved and the matrix (\ref{eq:15}) has been formed, then as initial trial coordinates, the points $x^q$, $0 \leq q \leq K$, can be used with values $z^q = F(\lambda,y^q)$. Moreover, the computation of the convolution (\ref{eq:05}) does not require the new calculation of criteria values at these points, because these values are taken from the matrix (\ref{eq:15}). After that, one can start optimization algorithm according to the Rules 1-5. Such the scheme reduces significantly the number of criteria computations when jointly solving the family of problems (\ref{eq:04}) with convolution (\ref{eq:05}) for different weight coefficients $\lambda$ \cite{ML_MCO_2023,Gergel2018,GergelKozinov2020}.

@@

For computation of $R_{ps}(i)$ we propose to apply the machine learning methods. It should be noted that in classical combinations of machine learning and optimization methods, the latter are used as an auxiliary internal techniques that improve the machine learning functioning. 
\textcolor[rgb]{0,0,1}{Other popular approaches use machine learning for auto-configuration (tuning the hyper-parameters) of optimization algorithms (see, e.g., \cite{Bezerra2020,Lopez2016} and the references given therein)}.
In our approach, on the contrary, machine learning methods are a part of the optimization algorithms enhancing their work.
